---
title: Going from hdf5 to ?
output: github_document
---

## Understand .h5

`/dataset1/data1/data` has what is in `dataset1/data1/what` which in our case is a 360x425 matrix explained here by Chris.

<div style="width:100%; background-color:#f00">
<table style="width:100%;">
<thead>
  <tr>
    <th></th>
    <th></th>
    <th colspan="5">Angles from due north</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td></td>
    <td></td>
    <td>1</td>
    <td>2</td>
    <td>3</td>
    <td>...</td>
    <td>360</td>
  </tr>
  <tr>
    <td rowspan="5">Distance <br>from radar</td>
    <td>1</td>
    <td colspan="5" rowspan="5">Values in the matrix give the values for the radar variable – dataset1/data1/what tells you what the variable is under “quantity”. dataset1/data1, dataset1/data2, etc, give the 10 different radar variables.</td>
  </tr>
  <tr>
    <td>2</td>
  </tr>
  <tr>
    <td>3</td>
  </tr>
  <tr>
    <td>...</td>
  </tr>
  <tr>
    <td>425</td>
  </tr>
</tbody>
</table>
</div>

> If you convert the matrix to a long format with the distance (called “range”) and angle (“theta”) then use some trigonometry (https://github.com/biodar/bdformats/commit/20502b63b6f252c26e1eddf4487b25379bc50a54) then you can plot the coordinates (I haven’t incorporated the height):

```{r chris-code}
library(rgdal)
library(rhdf5)

f <- "https://github.com/biodar/bdformats/releases/download/1/sample.h5"
if(!file.exists(basename(f))) {
  download.file(f, destfile = basename(f))
}
# Load in example dataset
testData<-h5read(basename(f),"/dataset1/data1/data")
str(testData)

# Turn matrix into long format, add angles and distances from radar to build polar coordinates
data.df<-data.frame(value=as.vector(testData),theta=rep(c(1:360),425),radius=rep(c(1:425),each=360))

# Trigonometry to calculate x and y (note that the scan is a single elevation, so treating it as flat for now)
data.df$x = data.df$radius*cos(data.df$theta)
data.df$y = data.df$radius*sin(data.df$theta)

# Where value is zero, assuming NoData
data.df$value[data.df$value==0]<-NA

# Quick plot - big dataset!
plot(data.df$x,data.df$y,cex=0.1)

# Make dataset smaller by removing points further than 50 units (about 5km, I think) from the radar
data.df.small<-subset(data.df,data.df$radius<50)

# Plot the smaller dataset
plot(data.df.small$x,data.df.small$y,cex=0.1)

# Apply the value offest to convert the integer values stored in the h5 dataset back to the radar variable
# The gain and offset are in the "what" part of the h5 data1 dataset
#   :gain = 0.1; // double
#   :offset = -32.0; // double
# I am not sure if I have applied them correctly here...
data.df.small$value<-data.df.small$value*0.1-32

# Visualising the new, smaller data with the values.
library(ggplot2)
sp<-ggplot(data.df.small, aes(x=x, y=y,color=value)) + geom_point() 
sp + scale_color_gradient(low="blue", high="red")
```
## hdf5r

```{r read}

f = "~/Documents/Research/BioDAR/Example_Met_Office_H5_Data_Share/202007170002_polar_pl_radar20b2_augzdr_lp.h5"

library(rhdf5)
df = h5ls(f)
class(df)
head(df)
d = h5read(f, "dataset1/data1/data")
class(d)
nrow(d)
length(d)
```
